Radial Pulsations in Polaris: A Secondary Science Application of Cherenkov Telescopes via Intensity Interferometry

### 2025-07-17 Phase_II/draft/GAN.tex
article
rearrange

\section{Generative Adversarial Networks}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

@@

\begin{figure}

@@

	\includegraphics[width=\linewidth]{fig/ellipse1612.jpg}
	\caption{Merged image, which includes the original and the sparsely sampled Fourier plane. It is exactly what the GAN receives. }
	\label{fig:GANinput}
\end{figure}
\begin{figure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_learning_rate_disc_loss.pdf}
		\caption{Discriminator loss for three different learning rates. }
		\label{fig:Plot_learning_rate_discloss}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_learning_rate_gen_total_loss.pdf}
		\caption{Total generator loss for different learning rates (eqn. (\ref{eq:total_gen_loss})).}
		\label{fig:Plot_learning_rate_genloss}
	\caption{Generator and discriminator losses for three different learning rates. The upper figure shows the total discriminator loss, and the lower figure shows the total generator loss. There is no significant difference, but the highest learning rate might be prone to outliers. }
	\label{fig:Plot_learning_rate_loss}
\end{figure}
\begin{figure}
	\begin{subfigure}{\linewidth}

@@

	\begin{subfigure}{\linewidth}

@@

\end{figure}
\begin{figure*}

@@

\begin{figure*}

>>>>

\section{Generative Adversarial Networks}
\begin{figure*}[hbt]
	\centering
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{fig/analysis/Plot_learning_rate_disc_loss.pdf}
		\caption{Discriminator loss for three different learning rates. }
		\label{fig:Plot_learning_rate_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{fig/analysis/Plot_learning_rate_gen_total_loss.pdf}
		\caption{Total generator loss for different learning rates (eqn. (\ref{eq:total_gen_loss})).}
		\label{fig:Plot_learning_rate_genloss}
	\end{subfigure}\hfill
	\caption{Generator and discriminator losses for three different learning rates. The left figure shows the total discriminator loss, and the right figure shows the total generator loss. There is no significant difference, but the highest learning rate might be prone to outliers. }
	\label{fig:Plot_learning_rate_loss}
\end{figure*}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

@@

\begin{figure*}[hbt]

@@

@@

@@

\begin{figure*}[hbt]

@@

\begin{figure*}[hbt]

### 2025-07-17 Phase_II/draft/II.tex
article
rearrange

@@

\subsection{The signal for II}\label{sec:signal}
As a simple example, let us consider a pair of IACTs pointed at a star. Suppose the two telescopes simultaneously measure the intensity of radiation $I_1(t)$ and $I_2(t)$, respectively. The signals from these detectors are cross-correlated and averaged over time, yielding the second order ($n=2$) correlation of these intensities as \citep[cf.][]{acciari2020optical, 2013APh....43..331D}
\begin{equation}
	g^{(2)}= \frac{\left\langle I_1(t) \cdot I_2(t + \tau) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} 
	\label{eqn:HBT}
\end{equation}
where $\tau$ is the time delay between the telescopes. For spatially coherent and randomly polarized light, eq.~(\ref{eqn:HBT}) reduces to the relation \citep[sometimes called the Siegert relation, see e.g.,][]{acciari2020optical}.
\begin{equation}
	g^{(2)} = 1 + \frac{\Delta f}{\Delta \nu} \abs{V_{12}}^2
	\label{eqn:HBT2}
\end{equation}
where $\Delta f$ is the electronic bandwidth of the photon detectors which measure the intensities.  Values of $\Delta\nu\sim 1\,\mathrm{THz}$ and $\Delta f \sim 1\,\mathrm{GHz}$ are typical of recent work.  In eq.~(\ref{eqn:HBT2}), $V_{12}$, referred to as the complex visibility function, is the Fourier transform of the source brightness distribution. For a uniform disk it is given by
\begin{equation}
 V_{12} = 2 \frac{J_1(\pi\theta_D b)}{(\pi\theta_D b)}
\label{eqn:absvisib}
\end{equation}
where $\theta_D$ is the angular diameter of the star and $b$ is the radial coordinate in the conventional interferometric $(u , v) = (x/\lambda, y/\lambda)$ plane, with $\lambda$ representing the optical wavelength of the filter used for observation. It is evident from eq.~(\ref{eqn:absvisib}) that $V_{12}$ contains information about the star's angular diameter. However, the phase information is lost since we measure only the absolute value $\vert V_{12} \vert^2$. In observational astronomy, the correlation is often expressed in terms of the normalized contrast, given by:
\begin{equation}
	c = \frac{\left\langle \left( I_1(t) - \left\langle I_1 \right\rangle \right) \cdot \left( I_2(t + \tau) - \left\langle I_2 \right\rangle \right) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} = g^{(2)} - 1
\end{equation}
where, $\left\langle I_1 \right\rangle$ and $\left\langle I_2 \right\rangle$ denote the mean intensities from the two telescopes. Therefore, the signal measured by the photon detector in II, operating with an electronic bandwidth $\Delta f$ within the optical bandwidth $\Delta {\mathrm {\nu}}$ of the observational radiation, is
\begin{equation}
	c = g^{(2)} - 1 = \frac{\Delta f}{\Delta \nu} \abs{V_{12}}^2
	\label{eq:signal}
\end{equation}
$\abs{V_{12}}^2$ is a function of baseline $d = \sqrt{u^2 + v^2}$ on observational plane. Consequently, the strength of the signal is enhanced by employing a large number of baselines.

\subsection{The Signal-to-Noise Ratio for II}
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/telescope.png}
	\caption{The telescope configuration with similar properties each used to simulate the signal for II.}
	\label{fig:teles}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/baseline.png}
	\caption{The tracking of baselines with four telescopes arranged in fig.~\ref{fig:teles} for one night of observation.}
	\label{fig:base}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig/ellipse/ellipse1612.jpg}
	\caption{This figure shows the simulated fast rotating star. The brightness is highest at the pole and there is gravitational darkening at the equator.}
	\label{fig:image}
\end{figure}
\begin{figure*}
	\centering
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft.jpg}
		\caption{The signal (absolute value of the Fourier transform of the source).}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_log.jpg}
		\caption{The logarithm of the signal.}
	\end{subfigure}
	\caption{Absolute value of the two-dimensional Fast Fourier Transform of Fig.~\ref{fig:image} measured by Intensity Interferometry, and observation of maximum $(u, v)$ plane with the assumed telescope separations can be completed with the help of Earth rotation.}
	\label{fig:ft}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_base.jpg}
		\caption{The signal as samples by the baselines.}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_log_base.jpg}
		\caption{The logarithmic signal as sampled by the baselines.}
	\end{subfigure}
	\caption{The left panel shows the absolute value of the two-dimensional Fast Fourier Transform of fig.~\ref{fig:image} measured by baselines shown in fig.~\ref{fig:teles}. The right panel shows the same on logarithmic scale}
	\label{fig:ft_base}
\end{figure*}
The primary purpose of IACTs is to study high-energy gamma rays (with energy $E\ \geq 30$ GeV) arriving from cosmic sources, entering the Earth's atmosphere, and initiating Cherenkov showers in the upper atmosphere due to multiple scattering. These telescopes feature an array of mirrors that focus light onto a set of photo-multiplier tubes \citep[PMTs, see e.g.,][]{aleksic2016major}. In the simulation model adopted here, we consider a set of four IACTs, each with similar properties. The positional configuration of these IACTs is shown in Fig.~\ref{fig:teles}. The optical signal directed to a PMT is filtered using a spectral filter with a chosen mean observational wavelength $\lambda$ and corresponding bandpass $\Delta \lambda$. The use of filters not only reduces background noise but also improves the signal quality and the efficiency of the PMTs. Filtering background skylight becomes even more significant in II observations, as these are carried out during full moon nights. It is important to note that the light from the stellar source is focused on the PMT during II observations.

@@

Intensity Interferometry counts the photons arriving at the telescopes from the stellar object. The correlation of these photon arrivals at the telescopes yields the squared visibility Eq.~\eqref{eq:angular_size_meas}, as explained in subsection~\ref{sec:signal}. Fig.\ref{fig:ft} shows the signal from the source shown in Fig.~\ref{fig:image} using II, displayed on both linear and logarithmic scales. In practice, only a small part of this information is available, because one has a finite number of baselines corresponding to the finite number $N_T$ of telescopes at our disposal and a limited observation schedule. Therefore, we have simulated the II observation of a sky source by four telescopes (correlated in pairs) over one night. Using this modest amount of signal from one night's observation, we have trained a cGAN to construct the image of the source.

>>>>

\begin{figure}[hbt]
	\centering
	\includegraphics[width=\linewidth]{fig/baseline.png}
	\caption{The tracking of baselines with four telescopes arranged in fig.~\ref{fig:teles} for one night of observation.}
	\label{fig:base}
\end{figure}
\begin{figure}[hbt]
	\centering
	\includegraphics[width=\linewidth]{fig/ellipse/ellipse1612.jpg}
	\caption{This figure shows the simulated fast rotating star. The brightness is highest at the pole and there is gravitational darkening at the equator.}
	\label{fig:image}
\end{figure}
\begin{figure*}[hbt]
	\centering
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft.jpg}
		\caption{The signal (absolute value of the Fourier transform of the source).}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_log.jpg}
		\caption{The logarithm of the signal.}
	\end{subfigure}
	\caption{Absolute value of the two-dimensional Fast Fourier Transform of Fig.~\ref{fig:image} measured by Intensity Interferometry, and observation of maximum $(u, v)$ plane with the assumed telescope separations can be completed with the help of Earth rotation.}
	\label{fig:ft}
\end{figure*}

@@

\begin{figure*}[hbt]
	\centering
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_base.jpg}
		\caption{The signal as samples by the baselines.}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_log_base.jpg}
		\caption{The logarithmic signal as sampled by the baselines.}
	\end{subfigure}
	\caption{The left panel shows the absolute value of the two-dimensional Fast Fourier Transform of fig.~\ref{fig:image} measured by baselines shown in fig.~\ref{fig:teles}. The right panel shows the same on logarithmic scale}
	\label{fig:ft_base}
\end{figure*}
\subsection{The signal for II}\label{sec:signal}
As a simple example, let us consider a pair of IACTs pointed at a star. Suppose the two telescopes simultaneously measure the intensity of radiation $I_1(t)$ and $I_2(t)$, respectively. The signals from these detectors are cross-correlated and averaged over time, yielding the second order ($n=2$) correlation of these intensities as \citep[cf.][]{acciari2020optical, 2013APh....43..331D}
\begin{equation}
	g^{(2)}= \frac{\left\langle I_1(t) \cdot I_2(t + \tau) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} 
	\label{eqn:HBT}
\end{equation}
where $\tau$ is the time delay between the telescopes. For spatially coherent and randomly polarized light, eq.~(\ref{eqn:HBT}) reduces to the relation \citep[sometimes called the Siegert relation, see e.g.,][]{acciari2020optical}.
\begin{equation}
	g^{(2)} = 1 + \frac{\Delta f}{\Delta \nu} \abs{V_{12}}^2
	\label{eqn:HBT2}
\end{equation}
where $\Delta f$ is the electronic bandwidth of the photon detectors which measure the intensities.  Values of $\Delta\nu\sim 1\,\mathrm{THz}$ and $\Delta f \sim 1\,\mathrm{GHz}$ are typical of recent work.  In eq.~(\ref{eqn:HBT2}), $V_{12}$, referred to as the complex visibility function, is the Fourier transform of the source brightness distribution. For a uniform disk it is given by
\begin{equation}
 V_{12} = 2 \frac{J_1(\pi\theta_D b)}{(\pi\theta_D b)}
\label{eqn:absvisib}
\end{equation}
where $\theta_D$ is the angular diameter of the star and $b$ is the radial coordinate in the conventional interferometric $(u , v) = (x/\lambda, y/\lambda)$ plane, with $\lambda$ representing the optical wavelength of the filter used for observation. It is evident from eq.~(\ref{eqn:absvisib}) that $V_{12}$ contains information about the star's angular diameter. However, the phase information is lost since we measure only the absolute value $\vert V_{12} \vert^2$. In observational astronomy, the correlation is often expressed in terms of the normalized contrast, given by:
\begin{equation}
	c = \frac{\left\langle \left( I_1(t) - \left\langle I_1 \right\rangle \right) \cdot \left( I_2(t + \tau) - \left\langle I_2 \right\rangle \right) \right\rangle}{\langle I_1(t) \rangle \cdot \langle I_2(t) \rangle} = g^{(2)} - 1
\end{equation}
where, $\left\langle I_1 \right\rangle$ and $\left\langle I_2 \right\rangle$ denote the mean intensities from the two telescopes. Therefore, the signal measured by the photon detector in II, operating with an electronic bandwidth $\Delta f$ within the optical bandwidth $\Delta {\mathrm {\nu}}$ of the observational radiation, is
\begin{equation}
	c = g^{(2)} - 1 = \frac{\Delta f}{\Delta \nu} \abs{V_{12}}^2
	\label{eq:signal}
\end{equation}
$\abs{V_{12}}^2$ is a function of baseline $d = \sqrt{u^2 + v^2}$ on observational plane. Consequently, the strength of the signal is enhanced by employing a large number of baselines.

\subsection{The Signal-to-Noise Ratio for II}
The primary purpose of IACTs is to study high-energy gamma rays (with energy $E\ \geq 30$ GeV) arriving from cosmic sources, entering the Earth's atmosphere, and initiating Cherenkov showers in the upper atmosphere due to multiple scattering. These telescopes feature an array of mirrors that focus light onto a set of photo-multiplier tubes \citep[PMTs, see e.g.,][]{aleksic2016major}. In the simulation model adopted here, we consider a set of four IACTs, each with similar properties. The positional configuration of these IACTs is shown in Fig.~\ref{fig:teles}. The optical signal directed to a PMT is filtered using a spectral filter with a chosen mean observational wavelength $\lambda$ and corresponding bandpass $\Delta \lambda$. The use of filters not only reduces background noise but also improves the signal quality and the efficiency of the PMTs. Filtering background skylight becomes even more significant in II observations, as these are carried out during full moon nights. It is important to note that the light from the stellar source is focused on the PMT during II observations.

@@

Intensity Interferometry counts the photons arriving at the telescopes from the stellar object. The correlation of these photon arrivals at the telescopes yields the squared visibility Eq.~\eqref{eq:angular_size_meas}, as explained in subsection~\ref{sec:signal}. Fig.\ref{fig:ft} shows the signal from the source shown in Fig.~\ref{fig:image} using II, displayed on both linear and logarithmic scales. In practice, only a small part of this information is available, because one has a finite number of baselines corresponding to the finite number $N_T$ of telescopes at our disposal and a limited observation schedule. Therefore, we have simulated the II observation of a sky source by four telescopes (correlated in pairs) over one night. Using this modest amount of signal from one night's observation, we have trained a cGAN to construct the image of the source.
\begin{figure}[hbt]
	\centering
	\includegraphics[width=\linewidth]{fig/ellipse1612.jpg}
	\caption{Merged image, which includes the original and the sparsely sampled Fourier plane. It is exactly what the GAN receives. }
	\label{fig:GANinput}
\end{figure}
